Visual narratives of the Covid-19 pandemic

### 2022-01-17 JDSSV_manuscript_knitr.Rnw
Adding in the spiral and lining up cases by wave

>>>>

There are many different ways that designers can use time-series data to provide additional contextual information and facilitate comparisons. \Cref{fig:ft-waves} shows covid cases, positivity rate, and hospital admissions from South Africa's Gauteng province, which was one of the first areas to experience a large wave of the Omicron variant. This chart provides context for the Omicron wave, showing that while infections are occurring at a much faster rate than in previous waves, hospital admissions seem to be approximately following previous trends.

\begin{figure}\centering
\includegraphics[width=\linewidth]{ft-waves}\\
\includegraphics[width=\linewidth]{ft-waves-linear}
\caption{Covid Cases, positivity rate, and hospital admissions from South Africa's Gauteng province, on a log scale (top) and linear scale (bottom). Successive waves are overlaid, showing that the increase in cases due to the Omicron variant was sharper than any previously measured increase, while hospitalizations were much more comparable to previous waves. Source: John Burn-Murdoch, \url{https://twitter.com/jburnmurdoch/status/1466480113487392769}}
\label{fig:ft-waves}
\end{figure}

While traditionally, time-series information has been limited to charts with time on the x-axis, there have been many attempts to make use of the capabilities of web graphics and animation in order to convey temporal information in less traditional time-series forms. One controversial chart appeared in the New York Times; the x-axis was wrapped around an Archimedean spiral to provide a sense of periodicity in the year-by-year evolution of case counts, as shown in \Cref{fig:nyt-spiral}. The original form of this chart is prone to the line-width illusion \citep{vanderplasSignsSineIllusion2015}, in addition to the well-known problems with polar charts \citep{hofmannGraphicalTestsPower2012,waldnerComparisonRadialLinear2020}. The re-envisioned version, which alignes counts on the outside of the spiral, and displays them as radial lines, mitigates some of the issues with the line-width illusion by directly showing each line as its own entity, facilitating direct comparisons. Still, it is more work to interpret and compare peaks on this chart than on a similar linear time series chart with the same data, as the reader must assess line width and then do a mental rotation and shift operation in order to compare with any other time period.

\begin{figure}\centering
\includegraphics[width=.45\linewidth]{nyt-spiral}\hfill\includegraphics[width=.45\linewidth]{spiral-rework}
\caption{Covid cases in the United States, 2020-2022. The original graph from the New York Times is on the left, and a reenvisioned version which is more perceptually friendly is on the right. Source: (Left) New York Times. Here's When We Expect Omicron to Peak \url{https://www.nytimes.com/2022/01/06/opinion/omicron-covid-us.html}, (Right) Covid Spiral, by Sourya Shrestha, \url{https://github.com/souryashrestha/covid_spiral}.}
\label{fig:nyt-spiral}
\end{figure}
